Hydra Battles in Coq

### 2020-10-30 doc/hydras.tex
correction of typos

This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:\gallina.

@@

The battle we are studying seems to be awfully long. Let us concentrate our
tests on some particular events : the states where $\texttt{nh}=0$

From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} will be
equal to zero.

@@

Thus, $ (1 + 11)$ rounds later, the \texttt{n1} field will be equal to $2$, and 
\texttt{nh}  will equal to $0$.

@@

\begin{Coqsrc}
Definition exp2 n := iterate (fun n => 2 * n) n 1.
\end{Coqsrc}
Using some facts (proven in \url{../src/html/BigBattle.html}),we get several  minorations.

@@

It is easy to represent the finite ordinals, successors and limits.

@@

Definition mult_fin  (alpha : t) (p : nat): t :=
  match alpha, p with
 |  (0,0), _  => zero
 |  _, 0 => zero
 |  (0, n), p => (0, n * p)
 |  ( n, b),  n' => ( n *  n', b)
               end.
Infix "*" := mult_fin : o2_scope.

@@

Definition fin_mult (n:nat)(beta : t) : t :=
  match n, beta with
 |  0, _  => zero
 |  _, (0,0) => zero
 |   n , (0,n') => (0, (n*n')%nat)
 |  n, (n',p') => (n', (n * p')%nat)
 end.

Example e1 : fin_mult 3 (omega * 7 + 15) = omega * 7 + 45.
Proof. reflexivity. Qed.

Example e2 :  omega * 2 + 2 + omega * 3 + 10 = omega * 5 + 10.
Proof. reflexivity. Qed.
\end{Coqsrc}

@@

Note that the type (\texttt{t 0}) is empty, and that \texttt{Program} generates an obligation
for verifying that the constraint $i<n$ is fullfilled when one tries to build the $i$-th ordinal inside the type (\texttt{t $n$}).

@@

@@

And we have now our transfinite recursor:
\index{Maths!Transfinite induction}

@@

\subsubsection{Definition of \texttt{oplus}}
\label{sec:orgheadline84}
\index{Functions!oplus @ oplus (Hessenberg commutative sum)}

@@

\begin{Coqsrc}
Corollary Cor12 (alpha : T1) :  nf alpha ->
         forall beta i n, beta  t1< alpha  ->
                (i < n)%nat ->
                 const_pathS i alpha beta ->
                 const_pathS n alpha beta.
Proof.
  transfinite_induction_lt alpha.
  (* (long) proof skipped *)
\end{Coqsrc}

@@

\begin{Coqsrc}
Corollary Cor12' (alpha : T1) :  nf alpha ->
         forall beta i n, (beta t1< alpha)%t1 ->
               (i <= n)%nat ->
               const_pathS i alpha beta ->
               const_pathS n alpha beta.
\end{Coqsrc}

@@

 \noindent
\emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
\begin{Coqsrc}
Example Ex1 : mlarge (omega * omega) (interval 6 510).
\end{Coqsrc}

@@

The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte says that this function is \emph{normal}).

>>>>

This chapter is dedicated to the representation of hydras and rules of the hydra game in \coq's specification language:~\gallina.

@@

The battle we are studying seems to be awfully long. Let us concentrate our
tests on some particular events : the states where $\texttt{nh}=0$.
From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.

@@

Thus, $ (1 + 11)$ rounds later, the \texttt{n1} field is equal to $2$, and 
\texttt{nh}   to $0$.

@@

\begin{Coqsrc}
Definition exp2 n := iterate (fun n => 2 * n) n 1.
\end{Coqsrc}
Using some facts (proven in 
\href{../src/html/hydras.Hydra.BigBattle.html}{hydras.Hydra.BigBattle}),we get several  minorations.

@@

It is easy to represent  finite ordinals, successors and limits.

@@

Definition mult_alpha_n  (alpha : t) (p : nat): t :=
  match alpha, p with
 |  (0,0), _  => zero
 |  _, 0 => zero
 |  (0, n), p => (0, n * p)
 |  ( n, b),  n' => ( n *  n', b)
               end.
Infix "*" := mult_alpha_n : o2_scope.

@@

Definition mult_n_alpha (n:nat)(alpha : t) : t :=
  match n, alpha with
 |  0, _  => zero
 |  _, (0,0) => zero
 |   n , (0,n') => (0, (n*n')%nat)
 |  n, (n',p') => (n', (n * p')%nat)
 end.

Example e1 : (omega * 7 + 15) * 3 = omega * 21 + 15.
Proof. reflexivity. Qed.

Example e2 :  mult_n_alpha 3 (omega * 7 + 15) = omega * 7 + 45.
Proof. reflexivity. Qed.
\end{Coqsrc}

@@

Note that the type (\texttt{t 0}) is empty, and that, for any natural number
 $n$, $n$ does not belong to (\texttt{t $n$}).

@@

\index{Projects}
\begin{project}
Consider two isomorphic ordinal notations \texttt{OA} and \texttt{OB}.
Prove that, if \texttt{OA} [resp. \texttt{OB}] is a correct implementation 
of $\alpha$ [resp. $\beta$], then $\alpha=\beta$.
\end{project}

@@

@@

\subsubsection{Definition of \texttt{oplus}}
\label{sec:orgheadline84}
%\index{Functions!oplus @ oplus (Hessenberg commutative sum)}

@@

\begin{Coqsrc}
Corollary Cor12 (alpha : T1) :  nf alpha ->
         forall beta i n, beta  t1< alpha  ->
                i < n ->
                 const_pathS i alpha beta ->
                 const_pathS n alpha beta.
Proof.
  transfinite_induction_lt alpha.
  (* (long) proof skipped *)
\end{Coqsrc}

@@

\begin{Coqsrc}
Corollary Cor12_1 (alpha : T1) :
  nf alpha ->
  forall beta i n, beta t1< alpha ->
                   i <= n ->
                   const_pathS i alpha beta ->
                   const_pathS n alpha beta.
\end{Coqsrc}

@@

\vspace{4pt}
 \noindent
\emph{From Module~\href{../src/html/hydras.Epsilon0.Large_Sets_Examples.html}{Epsilon0.Large\_Sets\_Examples}}
\begin{Coqsrc}
Example Ex1 : mlarge (omega * omega) (interval 6 510).
\end{Coqsrc}

@@

The ordering function $\phi_0$ of the set \texttt{AP} is defined on the full set $\mathbb{O}$ and is continuous (Schütte calls such a  function  \emph{normal}).
